---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RSExtraction

<!-- badges: start -->
<!-- badges: end -->

The goal of RSExtraction is to provide a resource to perform analysis presented in Lehner et al. (2024). Currently, this methodology is developed for DEM datasets projected in UTM. Other coordinate systems can be used, although the unit vector surfaces are unreliable due to the discrepancy between lat/long degrees vs meters between the planar extent and elevation values. Limited testing has shown that the ridge and swale features extracted do not appear to be greatly affected and seem reliable.

## Installation

You can install the development version of RSExtraction from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("jacoblehner/RSExtraction")
```
## Using the package:
```{r exampleUse, eval=FALSE, include=TRUE}
# Load libraries
library(RSExtraction)
library(terra)
library(viridis)

# Load your DEM using terra package
data <- rast("/path/to/DEM")

# Extract features
# set window size and if you would like to remove noise
morph <- UV.RS.extract(inRas = data, win = 3, rm.S = TRUE)

# Divide the results into the unit vector and ridge-swale components
uv.List <- morph$Unit.Vectors # unit vectors
rs.List <- morph$RS.Morph     # ridge-swale features

# Optional: divide further to separate the raster and matrix components
# unit vectors
uv.Mats <- uv.List$uv.Mats
uv.Rasts <- uv.List$uv.Rasts

# ridge-swale features
rs.Mats <- rs.List$rs.Mats
rs.Rasts <- rs.List$rs.Rasts

# The raster components are best for visualization and writing to file,
# the matrix components are included as an additional resource for simplifying
# further processing, if desired.

# Example of writing features to file
writeRaster(uv.Rasts, filename="/path/to/dir/rasName_UV.tif", overwrite = T)
writeRaster(rs.Rasts, filename="/path/to/dir/rasName_RS.tif", overwrite = T)
```

## Example 1: WGS 84 / UTM zone 32N (EPSG:32632) coord. reference
```{r example1}
library(RSExtraction)
library(terra)
library(viridis)

f <- system.file("ex/elev_vinschgau.tif", package="terra")
r <- rast(f) 
a <- disagg(r, 2.5)
data <- resample(r, a, "bilinear")
plot(data, col=turbo(200), legend = T)

# Extract features
morph <- UV.RS.extract(inRas = data, win = 3, rm.S = TRUE)

uv.List <- morph$Unit.Vectors # unit vectors
rs.List <- morph$RS.Morph     # ridge-swale features

# Visualize unit vector surfaces
plot(uv.List$uv.Rasts, col = gray(1:200/200), 
     legend = FALSE, smooth = TRUE)

# Visualize ridge and swale features
par(mfrow=c(1,2))
plot(uv.List$uv.Rasts$nX, col = gray(1:255/255),
     xlim = c(610000,650000), ylim = c(5145000,5190000),
     legend = FALSE, main = "Ridges.XY")
plot(uv.List$uv.Rasts$AvgEl, col = turbo(255, alpha = 0.5), legend = FALSE, add = TRUE)
plot(rs.List$rs.Rasts$r.XY, col = 'black', legend = FALSE, add = TRUE)

plot(uv.List$uv.Rasts$nX, col = gray(1:255/255),
     xlim = c(610000,650000), ylim = c(5145000,5190000),
     legend = FALSE, main = "Swales.XY")
plot(uv.List$uv.Rasts$AvgEl, col = turbo(255, alpha = 0.5), legend = FALSE, add = TRUE)
plot(rs.List$rs.Rasts$s.XY, col = 'black', legend = FALSE, add = TRUE)


```

## Example 2: lon/lat WGS 84 (EPSG:4326) coord. reference
```{r example2}
library(RSExtraction)
library(terra)
library(viridis)

f <- system.file("ex/elev.tif", package="terra")
r <- rast(f) 
a <- disagg(r, 3)
data <- resample(r, a, "bilinear")
plot(data, col=turbo(200), legend = T)

# Extract features
morph <- UV.RS.extract(inRas = data, win = 3, rm.S = TRUE)

uv.List <- morph$Unit.Vectors # unit vectors
rs.List <- morph$RS.Morph     # ridge-swale features

# Visualize unit vector surfaces
plot(uv.List$uv.Rasts, col = gray(1:200/200), 
     legend = FALSE, smooth = TRUE)

# Visualize ridge and swale features
par(mfrow=c(1,2))
plot(uv.List$uv.Rasts$nX, col = gray(1:255/255),
     legend = FALSE, main = "Ridges.XY")
plot(uv.List$uv.Rasts$AvgEl, col = turbo(255, alpha = 0.5), legend = FALSE, add = TRUE)
plot(rs.List$rs.Rasts$r.XY, col = 'black', legend = FALSE, add = TRUE)

plot(uv.List$uv.Rasts$nX, col = gray(1:255/255),
     legend = FALSE, main = "Swales.XY")
plot(uv.List$uv.Rasts$AvgEl, col = turbo(255, alpha = 0.5), legend = FALSE, add = TRUE)
plot(rs.List$rs.Rasts$s.XY, col = 'black', legend = FALSE, add = TRUE)


```
