---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RSExtraction

<!-- badges: start -->
<!-- badges: end -->

The goal of RSExtraction is to provide a resource to perform analysis presented in Lehner et al. (2024). Currently, this methodology is developed for DEM datasets projected in UTM. Other coordinate systems can be used, although the unit vector surfaces are unreliable due to the discrepancy between lat/long degrees vs meters between the planar extent and elevation values. Limited testing has shown that the ridge and swale features extracted do not appear to be greatly affected and seem reliable.

## Installation

You can install the development version of RSExtraction from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("jacoblehner/RSExtraction")
```

## Example 1: WGS 84 / UTM zone 32N (EPSG:32632) coord. reference
```{r example1}
library(RSExtraction)
library(terra)
library(viridis)

f <- system.file("ex/elev_vinschgau.tif", package="terra")
r <- rast(f) 
a <- disagg(r, 2.5)
data <- resample(r, a, "bilinear")
plot(data, col=turbo(200), legend = T)

# Extract features
morph <- UV.RS.extract(inRas = data, win = 3, rm.S = TRUE)

uv.List <- morph$Unit.Vectors # unit vectors
rs.List <- morph$RS.Morph     # ridge-swale features

# Visualize unit vector surfaces
plot(uv.List$uv.Rasts, col = gray(1:200/200), 
     legend = FALSE, smooth = TRUE)

# Visualize ridge and swale features
par(mfrow=c(1,2))
plot(uv.List$uv.Rasts$nX, col = gray(1:255/255),
     xlim = c(610000,650000), ylim = c(5145000,5190000),
     legend = FALSE, main = "Ridges.XY")
plot(uv.List$uv.Rasts$AvgEl, col = turbo(255, alpha = 0.5), legend = FALSE, add = TRUE)
plot(rs.List$rs.Rasts$r.XY, col = 'black', legend = FALSE, add = TRUE)

plot(uv.List$uv.Rasts$nX, col = gray(1:255/255),
     xlim = c(610000,650000), ylim = c(5145000,5190000),
     legend = FALSE, main = "Swales.XY")
plot(uv.List$uv.Rasts$AvgEl, col = turbo(255, alpha = 0.5), legend = FALSE, add = TRUE)
plot(rs.List$rs.Rasts$s.XY, col = 'black', legend = FALSE, add = TRUE)


```

## Example 2: lon/lat WGS 84 (EPSG:4326) coord. reference
```{r example2}
library(RSExtraction)
library(terra)
library(viridis)

f <- system.file("ex/elev.tif", package="terra")
r <- rast(f) 
a <- disagg(r, 5)
a <- disagg(r, 2.5)
data <- resample(r, a, "bilinear")
plot(data, col=turbo(200), legend = T)

# Extract features
morph <- UV.RS.extract(inRas = data, win = 3, rm.S = TRUE)

uv.List <- morph$Unit.Vectors # unit vectors
rs.List <- morph$RS.Morph     # ridge-swale features

# Visualize unit vector surfaces
plot(uv.List$uv.Rasts, col = gray(1:200/200), 
     legend = FALSE, smooth = TRUE)

# Visualize ridge and swale features
par(mfrow=c(1,2))
plot(uv.List$uv.Rasts$nX, col = gray(1:255/255),
     legend = FALSE, main = "Ridges.XY")
plot(uv.List$uv.Rasts$AvgEl, col = turbo(255, alpha = 0.5), legend = FALSE, add = TRUE)
plot(rs.List$rs.Rasts$r.XY, col = 'black', legend = FALSE, add = TRUE)

plot(uv.List$uv.Rasts$nX, col = gray(1:255/255),
     legend = FALSE, main = "Swales.XY")
plot(uv.List$uv.Rasts$AvgEl, col = turbo(255, alpha = 0.5), legend = FALSE, add = TRUE)
plot(rs.List$rs.Rasts$s.XY, col = 'black', legend = FALSE, add = TRUE)


```
